SAT Competition 2020\tnoteref{title}

### 2021-06-18 paper/main.tex
More...

One can see that the VBS solves all the instances that
were solved by at least one solver and solves each instance in the best observed time.
By quantifying how much each participating solver contributes
to the performance of the VBS, we may attempt to establish which
technology (as represented by the solvers) is the most important
(and to what degree) in the observed state of the art of SAT solving.
We consider here the following three related metrics conceptually derived
from the notion of VBS.

\begin{description}
\item[VBS-1]
``The fastest takes it all'':
For each solver we count the number of times, it was the fastest to solve an instance.
\item[VBS-2]
``Time aware, but proportional'':
A solver $S$ solving an instance $I$ in time $T^S_I$ accrues 
the following fraction of a point for solving $I$:
$T^\mathit{VBS}_I/T^S_I$, where $T^\mathit{VBS}_I$ is the solution time of the best solver on $I$.
\item[VBS-3]
``Split the point for solving'':
We award each solver $S$ solving an instance $I$ the fraction $1/|\mathcal{S}_I|$ of points, where
$\mathcal{S}_I$ is the set of solvers solving $I$.
\end{description}

We remark that the sum of VBS-1 points, as well as VBS-3 points, computed across all the solvers
is equal to the number of instances solved by at least one solver (later denoted \emph{total}).
This is obvious for VBS-1, as exactly one solver scores a point for solving an instance.
In the case of VBS-3, where we discard the information about the solution times, 
we evenly split the one point reward for solving an instance to those solvers which were successful on it.

In contrast, VBS-2 does not have this property as it in general distributes more than one point per instance.
Similarly to VBS-1, it takes the solution time into account,
and similarly to VBS-3 it does not award just the best solver on an instance.
For example, being twice as slow on an instance than the fastest solver results in a half point awarded.
And, if all the solvers solve an instance with the same solution time,
each solver scores a whole point for that instance.

@@

Table~\ref{tab:vbsMainALL} renders the just-described three metrics applied 
to the full results of the Main track. We can see that the respective leaderboards 
are generally dominated by \solver{Kissat} in at least one of its configurations.
VBS-1 tells us that \solver{Kissat-unsat} was most often the fastest solver,
in particular on \SI{11.5}{\percent} of all the solved instances.

The metric VBS-2 identifies as the best solver \solver{Kissat-sat}. Its leading score of \SI{32.9}{\percent} of the total
is a bit harder to interpret: A solver can score \SI{32.9}{\percent} of VBS-2 total
by solving \SI{32.9}{\percent} of the solved instances in the best observed time
and no others. But we know, thanks to its VBS-1, that \solver{Kissat-sat} solved \SI{9.8}{\percent}
of the solved instances in the best observed time (and some others).
A solver can also score \SI{32.9}{\percent} of VBS-2 total by solving all the solved instances,
but always being roughly 3 times slower than the VBS. 
\solver{Kissat-sat} lies (obviously) somewhere between these two extremes.

Finally, according to VBS-3, the best solver is \solver{Kissat} with 13.0 points,
which amounts to \SI{3.3}{\percent} of the distributed total score.
The VBS-3 metric is generally the most evenly distributed one,
at least among the first 10 solvers. 
(The last solver got 0.9 points which is \SI{0.221}{\percent} of the total.)
One can conclude from this that most of the benchmarks are solved
by most of the well performing solvers.

\subsection{Greedy Set Cover}
Another perspective on how much each solver contributes to the state of the art
can be obtained by attempting to construct a sequential schedule of solvers 
(rather than relying on an oracle to pick one for each instance, as with VBS)
and looking at how big a role each solvers plays in such a schedule. Since constructing 
an optimal schedule tends to be computationally hard, we start here by 
presenting a useful efficient alternative -- a greedy set cover approach.

With greedy set cover, we start with an empty schedule and iteratively consider 
each solver for the addition to the schedule obtained so far, picking the one
with the highest ``marginal contribution'' in terms of the number of problems
the new schedule will be able to solve. We demonstrate this on
the actual data from the competition, again taking the Main track results.

@@

A greedy set cover of the solved instances by solvers of the Main track is presented in Table~\ref{tab:gscMain}. 
In the first iteration, the solver which solved the highest number of instances is selected; in our case it was \solver{Kissat-sat} with 264 instances as we know already from Table~\ref{tab:results}.
Given those 264 instances as already covered, \solver{CaDiCaL-alluip} was the best in further contributing to the set by additional $22$ instances in iteration $2$.
We can see that the further iterations tend to add very little,
with the final four iterations adding one instance each. Note that each solver
that managed to solve an instance uniquely (i.e.~being the only solver to do so)
is showing up in the table. This is, for example the case of the last listed  
\solver{exp-V-LGB-MLD-CBT-DL}.
Highlighting unique solutions is a property of the greedy set cover metric
which we believe is congruent with documenting contributions to the state of the art.

@@

The greedy set cover technique focuses only on solving instances,
disregarding the time it would take to execute the obtained schedule. We can
also look at schedules that would fit in a prescribed time budget. A natural choice
of the budget seems to be the original time limit of \SI{5000}{\second}.

@@

By employing a brute-force approach, we first construct a sequences of schedules 
where the $i$-th schedule splits the available time of \SI{5000}{\second} uniformly among
$i$ solvers and solves the highest number of instances under these constraints.
The results are presented in Table~\ref{tab:MainUniformTimeSplit}.
We can see that the initial increase from 264 to 278 of ``covered'' instances 
when using two solvers instead of one (although allowing each to only use half of the time)
does not continue further with additional solvers allowed,
although its still better to use three solvers in a fair time split 
(and cover 272 instances) than just one.
One could use this observation to conjecture that the really hard 
problems that were solved actually require quite a large time to 
get ``cracked'' by any solver and thus the advantage of adding more solvers to the schedule
quickly diminishes.

@@

Finding a solution which satisfies all the hard constraints and as many soft constraints as possible,
Z3 comes up with the schedule shown in Table~\ref{tab:MainZ3Schedule} (in under two hours on a single core of a \SI{2.30}{\giga\hertz} CPU).
The table is sorted by $R^S$, the time the schedule allocates to individual solvers, and skips the zero entries.
It is not clear to what degree is the obtained schedule unique and how much it relies on each solver being present and for how long.
Nevertheless, it is interesting to observe the total number of problems covered, here 286,
and compare it to the 278 achieved in Table~\ref{tab:MainUniformTimeSplit} with the uniform split and two solvers.

As can be seen from the ``contributes'' column, the presence of \solver{f2trc-DL} in the schedule is not necessary.
% The following three lines could be skipped without much harm:
All the 15 instances this solver solves under 76 seconds were already covered by the preceding three solvers.
This result, however, does not signify a bug, as Z3 was not asked to produce a schedule with a minimal number of
participating solvers.
Indeed, allowing any other solver to run for the 76 ``wasted'' seconds would not increase the overall total.

@@

Interestingly, each of the first five tuples $T_{k \leq 5}$ contains exactly one of the three \solver{Kissat} variants. 
For $i < 5$ the relation $T_{i} \subset T_{i+1}$ holds only under projection to base solvers due to the fluctuating variants of \solver{Kissat}. 
The composition changes more strongly in $T_6$. 
Interestingly, we now have both variants $\{ \solver{Kissat-sat}, \solver{Kissat-unsat} \} \subset T_6$, and moreover it holds that $T_6 \subset T_7$.
All solvers in $T_{k \leq 7}$ belong the top performing solvers which received an award in the Main track, with the only exception being $\solver{Scavel01} \in T_4 \cap T_5$ (cf.~Table~\ref{tab:results}).

@@

Contributions to the VBS can be well captured by clustering the instance space into instance families. 
We evaluate the runtimes of the three winning solvers of the Main track on those new families which are represented by at least $14$ instances (cf.~Table~\ref{tab:families})
and report their place and score in Table~\ref{tab:eval-by-family}. 
Interestingly, the overall winning solver \solver{Kissat-sat} is outperformed by the second and third ranked solvers \solver{Relaxed-newTech} and \solver{CMS-ccnr-lsids} on the \emph{Anti-Bandwidth}, \emph{Vlsat}, and \emph{Influence Maximization} families by a large margin.

@@

% \begin{equation}
% \label{eq-similarity}
\[
	similarity(S, S') =\quad 1 - \frac{\sum{|S_{i} - S'_{i}|}}{316 \cdot \num{10000}}
\]
% \end{equation}
We calculate the \emph{similarity} of the $30$ solvers with the
best average PAR-2 score in the Main track. The results are displayed in
Figure~\ref{fig-similarity-main} as a \emph{heat map}, similar to the visualization in \cite{Balyo:2015:SATRace}. Additionally, the result
of hierarchically clustering the solvers based on their similarity is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example, enabling the \emph{trail
  saving} in \solver{CaDiCaL-alluip} has no impact on the runtime, resulting in
a similarity above \num{0.999}. Therefore the two solvers are joined low in the
dendrogram.
% A similar visualization was used for the results of the SAT Race 2015 \cite{Balyo:2015:SATRace}.

One big cluster that we identify consists of the \solver{Maple}-decedants, 
all of which are modifications of the winners in the SAT Competition 2018 and
the SAT Race 2019.
The similarity within the cluster is high, with the exception of
\solver{Scavel} and \solver{exp\_V\_MLD\_CBT\_DL}.
They form a sub-cluster
with a lower similarity to the rest, but a high similarity within. In fact, the highest measured similarity,
 % of $0.96$
besides the aforementioned \solver{CaDiCaL-alluip}, is observed between them.
The two solvers are both based on
\solver{MapleLCMDistChronoBT-dl-v2.2}, but have different authors.
The high similarity suggests that the
changes they made either result in a very similar behavior or do not have a
significant impact on the runtime performance.

The two configurations of \solver{Relaxed} by \emph{Zhang} and \emph{Cai}
use the same code-base as a lot of solvers in the \solver{Maple}-cluster, but their
overall performance is better and closer to the \solver{CMS}-cluster.
The three \solver{CMS} configurations
differ in their implementation of stochastic local search
(SLS) and have similar performance.
The 2020 version of \solver{CaDiCaL}
performs worse than the modifications based on the 2019 version and
does not quite fit into any cluster.

The last cluster to the left of the heat map is comprised of the other solvers
originally written by \emph{Biere}. Interesting to note is that the
\solver{Kissat} configuration specialized for unsatisfiable instances joins the
others in the cluster high in the dendrogram. In fact \solver{Kissat-unsat} has
the lowest average similarity to all other solvers in the top 30.
This suggests its importance for an optimal
portfolio.
% Average similarity
% | Kissat-unsat         |  21.75012381941168 |
% | Kissat               | 22.591060319234913 |
% | CaDiCaL-alluip       |  22.93588211785061 |
% | CaDiCaL-alluip-trail |  22.93631201417939 |
% | Kissat-sat           |  22.99738644487821 |

>>>>

One can see that the VBS solves all the instances that
were solved by at least one solver and solves each instance in the best observed time.
By quantifying how much each participating solver contributes
to the performance of the VBS, we may attempt to establish which
technology (as represented by the solvers) is the most important
(and to what degree) in the observed state of the art in SAT solving.
We consider here the following three related metrics conceptually derived
from the notion of VBS.

\begin{description}
\item[VBS-1]
``The fastest takes it all'':
For each solver, we count the number of times the solver was the fastest to solve an instance.
\item[VBS-2]
``Time aware, but proportional'':
A solver $S$ solving an instance $I$ in time $T^S_I$ accrues 
the following fraction of a point for solving $I$:
$T^\mathit{VBS}_I/T^S_I$, where $T^\mathit{VBS}_I$ is the runtime of the best solver on $I$.
\item[VBS-3]
``Split the point for solving'':
We award each solver $S$ solving an instance $I$ the fraction $1/|\mathcal{S}_I|$ of points, where
$\mathcal{S}_I$ is the set of solvers solving $I$.
\end{description}

We remark that the sum of VBS-1 points as well as the sum of VBS-3 points computed across all the solvers
is equal to the number of instances solved by at least one solver (later denoted \emph{total}).
This is obvious for VBS-1, as exactly one solver scores a point for solving an instance.
In the case of VBS-3, where we discard the information about the solution times, 
we evenly split the one point reward for solving an instance among those solvers which succeeded in solving the instance.
In contrast, VBS-2 does not have this property as it in general distributes more than one point per instance.
Similarly as VBS-1, it takes the solution time into account.
Similarly as VBS-3, it does not award just the best solver on an instance.
For example, a solver that uses twice as much runtime as  the fastest solver on an instance received half a point.
Further, if all the solvers solve an instance equally fast,
each solver receives a whole point for the instance.

@@

Table~\ref{tab:vbsMainALL} provides the result of applying the just-described three metrics 
to the full results of the Main track. We can see that the respective leaderboards 
are generally dominated by \solver{Kissat} in at least one of its configurations.
VBS-1 tells us that \solver{Kissat-unsat} was most often the fastest solver,
in particular on \SI{11.5}{\percent} of all the solved instances.

The metric VBS-2 identifies \solver{Kissat-sat} as the best solver. Its leading score of \SI{32.9}{\percent} of the total
is somewhat more difficult to interpret, though: a solver can score \SI{32.9}{\percent} of VBS-2 total
by solving \SI{32.9}{\percent} of the solved instances in the best observed time
and no others. Howeer,  we see from its VBS-1 score that \solver{Kissat-sat} solved \SI{9.8}{\percent}
of the solved instances in the best observed time (and some others).
A solver can also score \SI{32.9}{\percent} of VBS-2 total by solving all the solved instances,
but always being roughly three times slower than the VBS. 
The performance of \solver{Kissat-sat} lies (obviously) somewhere between these two extremes.

Finally, according to VBS-3, the best solver is \solver{Kissat} with 13.0 points,
which amounts to \SI{3.3}{\percent} of the distributed total score.
The VBS-3 metric is generally the most evenly distributed one,
at least among the first 10 solvers. 
(The last solver receives 0.9 points which is \SI{0.221}{\percent} of the total.)
One can conclude from this that most of the benchmarks are solved
by most of the well-performing solvers.

\subsection{Greedy Set Cover}
Another perspective on how much each solver contributes to the state of the art
can be obtained by attempting to construct a sequential schedule of solvers 
(rather than relying on an oracle to pick one solver for each instance, as with VBS)
and observing how big a role each solvers plays in such a schedule. Since constructing 
an optimal schedule tends to be computationally hard, we start here by 
presenting a a computationally more efficient alternative---a greedy set cover approach.

With greedy set cover, we start with an empty schedule and iteratively consider 
each solver for the addition to the schedule obtained so far, picking the one
with the highest ``marginal contribution'' in terms of the number of problems
the new schedule will be able to solve. We demonstrate this on
the actual data from the competition, again focusing in particular on the Main track results.

@@

A greedy set cover of the solved instances by solvers of the Main track is presented in Table~\ref{tab:gscMain}. 
In the first iteration, the solver which solved the highest number of instances is selected; in our case it was \solver{Kissat-sat} with 264 instances as we know already from Table~\ref{tab:results}.
With these
 264 instances already covered, \solver{CaDiCaL-alluip} is the best in further contributing to the set by additional $22$ instances in iteration $2$.
We can see that the further iterations tend to add very little,
with the final four iterations adding one instance each. Note that each solver
that managed to solve an instance uniquely (i.e., being the only solver that solved a particular instance)
showis up in the greedy set cover. 
% for example the case of the last listed  
%\solver{exp-V-LGB-MLD-CBT-DL}.
Highlighting solvers that uniquely solve instance is a property of the greedy set cover metric
which we believe is congruent with documenting contributions to the state of the art.
\textcolor{red}{congruent...? maybe a better choice of words here?}

@@

The greedy set cover 
disregards the time it would take to execute the obtained ``schedule'' of running the solvers covering instances. 
However, we can
also look at schedules that would fit in a prescribed time budget. A natural choice
of the budget seems to be the original time limit of \SI{5000}{\second}.

@@

To this end, by employing a brute-force approach, we first construct a sequences of schedules 
where the $i$-th schedule splits the available time of \SI{5000}{\second} uniformly among
$i$ solvers and solves the highest number of instances under these constraints.
The results are presented in Table~\ref{tab:MainUniformTimeSplit}.
We can see that the initial increase from 264 to 278 of ``covered'' instances 
when using two solvers instead of one (although allowing each to only use half of the time)
does not continue further with additional solvers allowed,
although its still better to use three solvers in a fair time split 
(and cover 272 instances) than just one.
Based on this observation, it is plausible  that the really hard 
instance that were solved actually may require quite large runtime to 
get ``cracked'' by any solver and thus the advantage of adding more solvers to the schedule
quickly diminishes.

@@

Finding a solution which satisfies all the hard constraints and as many soft constraints as possible,
Z3 provided  the schedule shown in Table~\ref{tab:MainZ3Schedule} (in under two hours on a single core of a \SI{2.30}{\giga\hertz} CPU).
The table is sorted by $R^S$, the time the schedule allocates to individual solvers, with zero entries ignored.
It is not clear to what degree is the obtained schedule unique and how much it relies on each solver being present and for how long.
Nevertheless, it is interesting to observe the total number of problems covered, here 286,
and compare it to the 278 achieved in Table~\ref{tab:MainUniformTimeSplit} with the uniform split and two solvers.

As can be seen from the ``contributes'' column, the presence of \solver{f2trc-DL} in the schedule is not necessary.
% The following three lines could be skipped without much harm:
All the 15 instances this solver solves under 76 seconds were already covered by the preceding three solvers.
This result is due to the fact that Z3 was not asked to produce a schedule with a minimal number of
participating solvers.
Indeed, allowing any other solver to run for the 76 ``wasted'' seconds would not increase the overall total.

@@

Interestingly, each of the first five tuples $T_{k \leq 5}$ contains exactly one of the three \solver{Kissat} variants. 
For $i < 5$ the relation $T_{i} \subset T_{i+1}$ holds only under projection to base solvers due to the fluctuating variants of \solver{Kissat}. 
The composition changes more strongly in $T_6$. 
Interestingly, we now have both variants $\{ \solver{Kissat-sat}, \solver{Kissat-unsat} \} \subset T_6$, and moreover it holds that $T_6 \subset T_7$.
All solvers in $T_{k \leq 7}$ are among the top-performing solvers which received awards in the Main track, with the only exception of $\solver{Scavel01} \in T_4 \cap T_5$ (cf.~Table~\ref{tab:results}).

@@

Contributions to the VBS can becaptured by clustering the instance space into instance families. 
We evaluate the runtimes of the three winning solvers of the Main track on those new families which are represented by at least $14$ instances (cf.~Table~\ref{tab:families})
and report their place and score in Table~\ref{tab:eval-by-family}. 
Interestingly, the overall best solver \solver{Kissat-sat} is outperformed by the second and third ranked solvers \solver{Relaxed-newTech} and \solver{CMS-ccnr-lsids} on the \emph{Anti-Bandwidth}, \emph{Vlsat}, and \emph{Influence Maximization} families by a large margin.

@@

% \begin{equation}
% \label{eq-similarity}
\[
	similarity(S, S') =\quad 1 - \frac{\sum{|S_{i} - S'_{i}|}}{316 \cdot \num{10000}}
\]
% \end{equation}
We calculate the \emph{similarity} of the $30$ solvers with the
best average PAR-2 score in the Main track. The results are shown in
Figure~\ref{fig-similarity-main} as a \emph{heat map}, similar to the visualization in \cite{Balyo:2015:SATRace}. Additionally, the result
of hierarchically clustering the solvers based on their similarity is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example, enabling the \emph{trail
  saving} in \solver{CaDiCaL-alluip} has no impact on the runtime, resulting in
a similarity above \num{0.999}. Therefore the two solvers are joined low in the
dendrogram.
% A similar visualization was used for the results of the SAT Race 2015 \cite{Balyo:2015:SATRace}.

Interestingly, one identifiable large cluster consists of the \solver{Maple}-decedants, 
all of which are modifications of the winners in the SAT Competition 2018 and
the SAT Race 2019.
The similarity within the cluster is high, with the exception of
\solver{Scavel} and \solver{exp\_V\_MLD\_CBT\_DL}.
They form a sub-cluster
with a lower similarity compared to the rest, but a high similarity within. In fact, the highest measured similarity,
 % of $0.96$
besides the aforementioned \solver{CaDiCaL-alluip}, is observed between them.
The two solvers are both based on
\solver{MapleLCMDistChronoBT-dl-v2.2}, but have different authors.
This high similarity suggests that the
changes they made either result in a very similar behavior or do not have a
significant impact on the runtime performance.

The two configurations of \solver{Relaxed} by \emph{Zhang} and \emph{Cai}
use the same code-base as a lot of solvers in the \solver{Maple}-cluster. However, the
overall performance of the solver is better and closer to the \solver{CMS}-cluster.
The three \solver{CMS} configurations
differ in their implementation of stochastic local search
(SLS) and have similar performance.
The 2020 version of \solver{CaDiCaL}
exhibit weaker performance than the modifications based on the 2019 version and
does not quite fit into any cluster.

The leftmost cluster in the heat map is comprised of the other solvers
originally written by \emph{Biere}. What is interesting to note is that the
\solver{Kissat} configuration specialized for unsatisfiable instances joins the
others configurations in the cluster high in the dendrogram. In fact \solver{Kissat-unsat} has
the lowest average similarity to all other solvers in the top 30.
This suggests its importance for an optimal
portfolio.
% Average similarity
% | Kissat-unsat         |  21.75012381941168 |
% | Kissat               | 22.591060319234913 |
% | CaDiCaL-alluip       |  22.93588211785061 |
% | CaDiCaL-alluip-trail |  22.93631201417939 |
% | Kissat-sat           |  22.99738644487821 |
